Selective enhancement of an exploited species undermines community stability

### 2021-11-30 rmd/manuscript.Rmd
edit: manuscript

We use a multispecies Ricker model to simulate community dynamics with selective enhancement of a particular species. Specifically, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

Our model predicted a destabilizing effect of selective enhancement on ecological communities. With increasing released individuals $R$, communities showed greater temporal fluctuations as illustrated by increased CV (Figure XX). Since CV is a ratio of SD to mean, the increase of CV can be driven by a decreased mean, an increased SD, or both. In our case, the group-level analysis depicted a major influence of the decreased means of enhanced and unenhanced species groups (Figure XX). For enhanced species, intraspecific competition is responsible for the decreased mean. That is, released individuals induce strong overcompensation, where the negative effect of reduced per-capita reproduction exceeds the numerical benefit (increased reproducing individuals) of released individuals. Meanwhile, the decreased mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. This negative effect of released individuals through competition was noticeable even with weak interspecific competition that would otherwise allow stable coexistence. Combined, the total community density decreased more sharply than individual species groups with increased numbers of released individuals (XX). The SDs showed a similar trend, but the relationship was flatter than the means. CONCLUDING SENTENCE

ADDITIONAL PARAGRAPH ABOUT EXTERNAL CONTROL OF RELEASED INDIVIDUALS

>>>>

We use a multispecies Ricker model to simulate community dynamics with selective enhancement of a constituent species. Specifically, population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

@@

Our model predicted a destabilizing effect of selective enhancement on ecological communities. Increasing numbers of released individuals $R$ led to greater temporal fluctuations of communities as illustrated by increased CV (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals (increased reproducing individuals). Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. This negative effect of released individuals through competition was noticeable even with weak interspecific competition that would otherwise allow stable coexistence. Combined, the long-term average of the total community density decreased more sharply than individual species groups with increased numbers of released individuals (XX). The SDs showed a similar trend, but the relationship was flatter than the mean at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV.

The phenomena of the decreased long-term means should arise as long as humans control the number of released individuals (i.e., donor-control). The released individuals are free from any density dependent feedback in natural communities and therefore undermine the internal regulation of self-sustaining communities.

These results were qualitatively consistent across ecological scenarios (XX).

The desta
